---
title: "data wrangling ex2"
author: "sasiratkitt"
date: "August 25, 2018"
output: html_document
---

```{r load package, message = FALSE}
library(tidyverse)
library(dummies)
library(funModeling)
library(mice)
library(Hmisc)
```

### 0: Load the data in RStudio
After loading the dataset, I tried applying different techniques to identify missing values, which includes "" and NA values.

```{r load data, echo= TRUE}
setwd("~/R Projects/springboard_datascience_r/data_wrangling")
titanic0 <- read.csv("titanic3_original.csv", na.strings = TRUE, stringsAsFactors = FALSE)
tbl_df(titanic0)

#Checking NA, zeros, data type and unique values with various functions. However, this function fails to detect "" values.
my_data_status <- funModeling::df_status(titanic0)

#colSums function can also be used in this case
colSums(is.na(titanic0))

colSums(titanic0 == "")

#Subseting
#test <- titanic0[titanic0$embarked == "C",]
#count(titanic0[titanic0$embarked == "C",])
```

### 1: Port of embarkation
```{r embarked}
titanic <- titanic0
titanic <- titanic %>% mutate(embarked = ifelse(embarked == "", "S", embarked))

#Question: How do we write lapply code for this case
#titanic$embarked<- lapply(titanic$embarked, replace(titanic$embarked, "", "S"))
```

### 2: Age
To replace missing 'age' values, we can use mean or median as a simplified method. We also cross-checked at how average age varies by passenger gender and observe distribution of age range

```{r age}
#replace missing 'age' with mean or median of data rows without NA
mean(titanic$age, na.rm = TRUE)
median(titanic$age, na.rm = TRUE)

titanic <- titanic %>% mutate(age = ifelse(is.na(age) | age == "" , mean(titanic$age, na.rm = TRUE), age))

#titanic$age <- impute(titanic$age, mean)

#exploring further, we look at how average age varies by passenger gender and observe distribution of age range.
hist(titanic0$age[titanic0$sex == "female"])
mean(titanic0$age[titanic0$sex == "female"], na.rm = TRUE)

hist(titanic0$age[titanic0$sex == "male"])
mean(titanic0$age[titanic0$sex == "male"], na.rm = TRUE)
```
Since there is not much difference in female vs. male average age or age distribution, average value of all passengers will be used.


### 3: Lifeboat
```{r lifeboat}
titanic$boat[titanic$boat == ""] <- NA

#Check that titanic$boat's missing values are replaced with NA
colSums(titanic == "")
```

### 4: Cabin
```{r cabin}
#77% of cabin information is missing. So it may not make sense to fill in the missing values
titanic <- titanic %>% mutate(has_cabin_number = ifelse(cabin == "", 0, 1))

#test correlation between passenger survival to cabin number ownership
cor(titanic$survived[!is.na(titanic$survived)], titanic$has_cabin_number[!is.na(titanic$survived)])
```

### 5: Submit the project on Github
```{r write}
write.csv(titanic, "titanic_clean.csv", row.names = FALSE)
```


